---
title: "TP2_VECM"
author: "GUENOUNE Loubna"
date: "2022-12-25"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing the data :

```{r}
library(readxl)
data <- read_excel("C:/Users/Lenovo/Downloads/April_data_6series.xls")

```

#A.Creating ts objects :
The EU time serie measures the era swap cryptocurrency change rate to us  dollars while,the GU measures the nasdaq stock index (in us dollar ).


```{r}
EU<- ts(log(data$es))
GU<- ts(log(data$nq))
```

Plotting the time series in the same graph :

```{r}
ts.plot(EU, GU, gpars = list(col = c("black", "red")))
```
Though the difference is big between,we can see they have similar movement and it is even better shown when plotting the time series individually and comparing the graphs :

```{r}
plot(EU)
plot(GU)

```

#Q.b : ERS unit root test

The whole VECM stands on the non stationnarity of the primary data, so we preform an ERS Test on both of the time series :

```{r}
library(urca)
EU.urers1 <- ur.ers(EU, type="P-test", model="trend")
summary(EU.urers1)
GU.urers2 <- ur.ers(GU, type="P-test", model="trend")
summary(GU.urers2)
```
the test statistic is clearly larger than critical values at 1 and 5 percent level of confidence for both series, so we reject the null hypothesis of no unit root 

we check for i(1):

# reject Null of no unit root -> exist unit root -> difference

we differentiate the time series and recheck for unit roots :


```{r}
dEU <- diff(EU)
dGU <- diff(GU)
dEU.urers1 <- ur.ers(dEU, type="P-test", model="trend")
summary(dEU.urers1)
dGU.urers2 <- ur.ers(dGU, type="P-test", model="trend")
summary(dGU.urers2)
par(mfrow=c(1,1))

plot(dEU, type='l', main="EURO TO USD vs. GBP TO USD ", col="blue")
lines(dGU, col="red")
```

The test statistic is smaller than the critical values at 1 5 and 10 percent , so we accept the null hypothesis of no unit root and both of the series are I(1).
The graph of differentiated time series shows a near identicalness of the time series.

#Q.c

Binding the two series in one vector object :

```{r}
library(lmtest)
library(sandwich)
library(MASS)
library(urca)
library(vars)
y <- cbind(EU,GU)
colnames(y) <- c("EU","GU")
y <- na.trim(y)
```


# determination of the number of lags to be included in cointegration test and in VEC model

```{r}
y.VAR.IC <- VARselect(y, type="const")
nlags<-y.VAR.IC$selection["SC(n)"]
nlags

```

the schwarz criterion shows that the ideal number of lags to be included is 2.

```{r}
y <- window(y, start=160, end=200)
```

#cointegration test

we use the Johansen test since it's a bivariate model:

```{r}
library(urca)
y.CA <- ca.jo(y, ecdet="const", type="trace", K=nlags, spec="transitory")
summary(y.CA)
y.CA <- ca.jo(y, ecdet="const", type="eigen", K=nlags, spec="transitory")
summary(y.CA)
lttest(y.CA, r=1)
```

The results of Johansen’s Trace test show that we can reject the null hypothesis of r=0 at a 10 percent level, Therefore we can conclude that ERA Swap and Nasdaq stock index are cointegrated.
The results of maximum eigenvalue test also show that we can reject the null hypothesis. We can get same conclusion with above.
On the other hand, the result of LR test which tests for the presence of the constant term in the cointegration relationship suggests case 2(restricted constant).

# Q.d

# estimatation of  unrestriced VEC model :
```{r}
y.VEC <- cajorls(y.CA, r=1)
y.VEC
```
The cointegrating vector is estimated ββ = (1.000000,-0.4196125, -4.1470515). We have included a constant in the cointegrating relationship.
Adjustment parameters are estimated as αα = (-0.6465,0.8205).

#Q.E

# to see t-statistics and p-values

both α1α1 and α2α2 are statistically significant at 0.001 level. Adjustment
parameters α1α1 and α2α2 determine the speed of return to long run equilibrium. For long run
relationship to be stable, we need α1α1 < 0 or =0, α2α2 > 0 or =0 and at least one of them can not be equal to 0. In this result, α1α1 <0 and α2α2 <0, so it does not satisfy the condition for long run stable relationship.

```{r}
summary(y.VEC$rlm)
```

#Q.F

# testing  for restricted adjustment parameters alpha :

```{r}
rest.alpha <- matrix(c(1,0), c(2,1))
y.CA.ralpha <- alrtest(y.CA, A=rest.alpha, r=1)
summary(y.CA.ralpha)
```

this code is for the cointegraction vector wich helps us to determinate le long term relation between the variables in the system.

# Q.G
# vec2varX supports restricted VEC models:
```{r}
library(vars)
y.VAR <- vec2var(y.CA, r=1)
y.VAR.rest <-vec2var(y.CA, r=1)
y.VAR.rest
```

If α1α1 >0 and α2α2 <0 then EU would be growing and GU would be declining, taking the system even further away from the long run equilibrium. So we need to restrict one of them to
be zero. if α2α2 =0 and α1α1 is not zero, then y2,ty2,t, GU is a pure random walk and all the adjustment occurs in y1,ty1,t,EU

# forecast using VAR in levels

```{r}
y.VAR.fcst <- predict(y.VAR.rest, n.ahead=76, ci=0.95)
par( mar=c(4,4,2,1), cex=0.75)
plot(y.VAR.fcst)
```

The plot below shows the forecast using multistep scheme.


#Q.I
```{r}
first.m <- 29285
last.m <-29345
dEU.p1 <- window(dEU, end=last.m)
dGU.p2 <- window(dGU, start=first.m)
library(forecast)
par(mfrow=c(1,2))
Acf(dEU.p1, type="correlation", lag=48, main="ACF for EU")
Acf(dGU.p2, type="partial", lag=48, main="PACF for GU")
arma26 <- Arima(dGU.p2, order=c(2, 0, 6))
tsdiag(arma26, gof.lag=36)
```


 
 
 